---
title: "Tutorial for the implemented penalized approach for anomaly detection"
author: "AMVA4NewPhysics authors"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In the tutorial, we present how to use package PenalizedAD that contains the implementation of penalized model-based clustering approach and as well the following anomaly detection method.

First, let us load an example of the background data produced by a mixture of 3 Gaussian components.

```{r}
require(PenalizedAD)
data(bcg_data)
```

Given such the data we can use MAES algorithm for model-based clustering to perform classification.

```{r}
f <- MAES(bcg_data, 2, attr(bcg_data, "label"))
```

and plot the classificaion results
```{r}
plot(f[[8]],col=attr(bcg_data, "label"),
     main="Classification performance",ylab="Predicted classes", xlab = "Observations")
```

<!-- For the anomaly detection we make use of the above approach for background density estimation. We can load an example data and fit the anomaly model. -->
<!-- ```{r} -->
<!-- data(bcg_data) -->
<!-- data(experiment_data) -->

<!-- #1 fit the background for the warm start -->
<!-- E_bcg=MAES(bcg_data,2,attr(bcg_data, "label")) -->

<!-- #fit the signal -->
<!-- #potential model for experimental data if no significant signal components is found -->
<!-- sig <- E_bcg -->

<!-- #Iterate between fiting the bacground on data and fitting the signal on data2 -->
<!-- minbic <- Inf -->
<!-- n_iter <- 20 -->
<!-- sig <- signal(E_bcg, experiment_data,sig,gamma_eigen = 1, -->
<!--             gamma_mean = 4,n_iter,minbic,bcg_data) -->
<!-- ``` -->

