---
title: "Tutorial for the implemented penalized approach for anomaly detection"
author: "AMVA4NewPhysics authors"
date: "`r Sys.Date()`"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In the tutorial, we present how to use package PenalizedAD that contains the implementation of penalized model-based clustering approach and as well the following anomaly detection method.

First, let us load an example of the background data produced by a mixture of 3 Gaussian components.

```{r}
require(PenalizedAD)
data(bcg_data)
```

Given such the data we can use MAES algorithm for model-based clustering to perform classification.

```{r}
f <- MAES(bcg_data, 2, attr(bcg_data, "label"))
```

and plot the classificaion results
```{r}
plot(f[[8]],col=attr(bcg_data, "label"),
     main="Classification performance",ylab="Predicted classes", xlab = "Observations")
```

For the anomaly detection, we make use of the above approach for background density estimation. We can load an example data and fit the anomaly model.
```{r}
data(bcg_data)
data(experiment_data)

#1 fit the background as the potential background model if no signal is to be found
E_bcg=MAES(bcg_data,2,attr(bcg_data, "label"))

#fit the signal

#Iterate between fiting the bacground on data and fitting the signal on data2
minbic <- Inf
sig <- PAD(E_bcg, experiment_data,gamma_eigen = 1,
            gamma_mean = 4,minbic,bcg_data)

#The component means
sig[[1]]
```

